---
title: 'STAT 341/641 Lab: Week Six'
author: "Vera Yang"
date: "March 8, 2020"
output:
  pdf_document: default
  html_document: default
---
---

**STAT 341/641:**        Intro to EDA and Statistical Computing  
**Lab #5:**              Rejection Sampling and the Bootstrap  
**Teaching Assistant:**  Yanjun Liu

* * *

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Directions:**  You will use the lab time to complete this assignment.

* * *

#**Task: Practice rejection sampling and the bootstrap**  


##1: Suppose we would like to obtain samples from a probability density that satisfies
$$f(x) \propto \exp\left\{-x^2/2 \right\}(\sin(6x)^2 + 3\cos(x)^2).$$
The symbol $\propto$ indicates that the density is proportional to $\exp\left\{-x^2/2 \right\}(\sin(6x)^2 + 3\cos(x)^2)$.  In general, two quantities $a$ and $b$ are proportional if $a = kb$ where $k$ is called the constant of proportionality.  For probability distributions, we frequently know the numerator, but not the denominator.  

Anyway, find a number $M \geq 1$ such that the envelope condition holds for proposal distribution $g(x) = N(0,1).$  Recall the envelope condition is $f(x) \leq M g(x)$ for all $x$.  

As a hint, look at the code chunk below.  Try looping over values for $M$.  Choose the smallest value of $M$ that satisfies the condition. 


**Solution:** The smallest M value is 10.
```{r}

myseq <- seq(from = -4, to = 4, by = .1)
f <- function(x){
  exp(-x^2/2)*(sin(6*x)^2 + 3*cos(x)^2)
}

#all(f(myseq) < M * dnorm(myseq))

# values for M will be 1 to 20
M <- c(1:20)

# this stores the minimum M value
min_M <- numeric(length(M))

for(i in 1:length(M)){
  if(all(f(myseq) < M[i] * dnorm(myseq))){
    min_M[i] <- M[i]
  #  print(paste0("M is", M[i]))
  }
}

min_M <- min_M[min_M>0]
min_M <- min(min_M)
min_M
```

##2: Write a loop to obtain 1,000 samples from $f(\cdot)$.  How many iterations does it take to get 1,000 samples?  
$f(\cdot)$
Edit the code chunk supplied below to accomplish this.

**Solution:**  It takes 1772 iterations to get 1,000 samples.
```{r}
set.seed(123)
mynum <- 0
N <- 1000
fsamples <- NULL
M <- 10 # this M value is from Question 1

# this variables keeps track the total number of iterations
# to obtain 1000 samples
count <- 0

while(mynum < N){
  ## sample from the proposal
  mysamp <- rnorm(1)
  
  ## compute r
  myr = (f(mysamp)) / 
    (M * dnorm(mysamp))

  if(sample(c(0,1), 1, prob = c(1-myr, myr)) == 1){
    ## record the sample
    fsamples <- c(fsamples, mysamp) 
    #plug this fsamples into the x inside Mg(x) for Question 3
    
    #mynum only tracks the number of samples that I'm keeping 
    # (ones that satisfy the condition)
    mynum <- mynum + 1 
  }

  # incease count by 1 after each iteration 
  # to find the total number of iterations
  count = count + 1
}

count

```

##3: Plot histograms of 1,000 samples from $Mg(x)$ and $f(x)$.  Use pastel colors with a low alpha (high level of transparency) in order to overlay the two distributions on the same plot.

**Solution:**  Blue is samples from Mg(x) and red is samples from f(x)!
```{r}
library(tidyverse)

# g(x) is dnorm(fsamples)
M <- 10
mgsamples <- M*dnorm(fsamples)  #this is Mg(x)

fx <- f(fsamples)  #this is f(x)

fsamples.df <- as.data.frame(fx)
mgsamples.df <- as.data.frame(mgsamples)

fg <- as.data.frame(cbind(fsamples.df, mgsamples.df))

# blue is samples from f(x)
# red is samples from Mg(x)
ggplot(data = fg) +
  geom_histogram(aes(fx), fill = "#FF9999", alpha = .5) +
  geom_histogram(aes(x = mgsamples), fill = "#56B4E9", alpha = .5) +
  ggtitle("Histogram of f(x) and Mg(x)") +
  xlab("f(x) and Mg(x)")
```

##4: Load the rousseeuw_outlier_set1.csv data set.  You are going to edit the code chunk below to create some interesting plots. For each bootstrap replication, record whether the first row of the data set has been chosen.  Then make two different plots of the regression lines:  one for all bootstrap samples with the first row and another for all bootstrap samples without the first row.  How many of the 1,000 bootstrap replications contain the first row?

**Solution:**  
```{r}
set.seed(641)
outs <- read.csv(file = "rousseeuw_outlier_set1.csv")
R <- 1000


# count the number times the first row is selected in every sample
count_first <- 0  

# matrix to record the regression coefficients
# for samples with the first row and without first row
with_first_betas <- matrix(0,R,2)
without_first_betas <- matrix(0,R,2)

# loop to get 1000 samples
for (j in c(1:R)){
  
  ## sample indices to make bootstrap sample
  inds <- sample(c(1:nrow(outs)), size = nrow(outs),replace = T)
  inds.df <- as.data.frame(inds)
  
  # count the number of times that first row is selected in a sample
  for(k in 1:nrow(inds.df)){
    if(inds.df[k,] == 1){
      count_first = count_first + 1
    }
  }
  
  # in each sample, if first row is selected,
  # compute the regression and store coefficients in with_first_betas
  if(count_first > 0){
    with_first.lm <- lm(Y ~ X, data = outs[inds,])
    with_first_betas[j,] <- coef(with_first.lm)
  }else{
    
    # else compute regression and store coefficients in without_first_betas
    without_first.lm <- lm(Y ~ X, data = outs[inds,])
    without_first_betas[j,] <- coef(without_first.lm)
  }
  
  # reset count for number of first row in each iteration
  count_first = 0
}

# remove the empty rows 
with_first_betas <- with_first_betas[with_first_betas[, 1] > 0,]
without_first_betas <- without_first_betas[without_first_betas[, 1] > 0, ]

# get the column means
with_first_mean <- colMeans(with_first_betas)
without_first_mean <- colMeans(without_first_betas)

# assign a colour to every set of regression coefficients
with_first_cols <- rainbow(nrow(with_first_betas), alpha = .20)
without_first_cols <- rainbow(nrow(without_first_betas), alpha = .20)

# plot the lines
plot(outs,typ="n", xlab="input", ylab="output",
     main = "Regression for Samples with First Row")
for (i in c(1:nrow(with_first_betas))){
  abline(a = with_first_betas[i,1], b = with_first_betas[i,2], 
         col=with_first_cols[i], lwd = .5)
}

plot(outs,typ="n", xlab="input", ylab="output",
     main = "Regression for Samples without First Row")
for (i in c(1:nrow(without_first_betas))){
  abline(a = without_first_betas[i,1], b = without_first_betas[i,2], 
         col=without_first_cols[i], lwd = .5)
}

```

```{r, include=FALSE}
# code given in lab
mybetas <- matrix(0,R,2)
betas <- colMeans(mybetas)
mycols <- rainbow(R,alpha = .20)
plot(outs,typ="n",xlab="input",ylab="output",main = "Regression with Outliers")
points(outs,pch = 20, col="blue")
#abline(res, col = "red")
for (j in c(1:R)){
  abline(a = mybetas[j,1], b = mybetas[j,2], col=mycols[j], lwd = .5)
}
```


* * *
